---
execute: 
  echo: false
---
## Introduction
As we all know that pandemic which was caused by wide spread of COVID virus started spring of 2020 (it was declared by WHO on 3/11/2020) and ended at spring of 2023 (declared by WHO that it ended on 5/11/2023).  During the pandemic, 
it was noticed that number of people who were infected by flu virus was much lower than prior years.  The hypothesis which I would like to propose and decided to choose as the data science project is whether if there is any correlation between COVID and Flu infection, so I searched for the data set of infection rate of both viruses that could be used to draw a graph and it turned out that they were available at CDC website.  Since COVID and Flu infection rate data were accessible separately and the data format was different, I had to do some data transformations in order to combine them into one data source and use it to draw on the same graphes.

## Line Graph of Infection Rate of COVID and Flu
The first graph is infection rate of COVID and Flu virus over time span of past few years including years we had pandemic.

```{r}
#| label: setup
#| message: false
#| warning: false
library(tidyverse)
library(dplyr)
suppressPackageStartupMessages(library(ISOweek))
```

```{r}
#| message: false
#| warning: false
get_monday_by_date <- function(date) {
  # Ensure the input is a date
  date <- as.Date(date)
  # Calculate the difference from Monday
  diff_to_monday <- wday(date, week_start = 1) - 1
  # Subtract the difference from the date
  monday_date <- date - diff_to_monday
  return(monday_date)
}
get_monday_by_year_week <- function(year, week) {
  week_in_string <- sprintf("%d-W%02d-1", year, week)
  monday_of_week <- ISOweek::ISOweek2date(week_in_string)
  return(monday_of_week)
}

covid_infection_rate <- 
  read_csv("data/weekly_covid_infection_rate.csv", col_types = cols(Date = col_date("%b %d %Y")), show_col_types = FALSE) |> filter(!is.na(Weekly_Percent_Test_Positivity))|> 
  mutate(TYPE = "covid", INFECTION_RATE = as.numeric(Weekly_Percent_Test_Positivity), REPORT_DATE = get_monday_by_date(Date))

flu_infection_rate <- read_csv("data/weekly_flu_infection_rate.csv", show_col_types = FALSE) |> mutate(TYPE = "flu", INFECTION_RATE = PERCENT_POSITIVE, REPORT_DATE = get_monday_by_year_week(YEAR, WEEK))

infection_rate <- bind_rows(covid_infection_rate |> select(INFECTION_RATE, TYPE, REPORT_DATE), flu_infection_rate |> select(INFECTION_RATE, TYPE, REPORT_DATE))




infection_rate |> ggplot(aes(x = REPORT_DATE, y = 
                               INFECTION_RATE, color = TYPE)) + geom_line(linewidth = 1) + 
  scale_x_date(
    date_breaks = "1 week"
  ) +labs(title = "Infection Rate of COVID and Flu (Type A and Type B combined)", subtitle = "Noticeable Negative Correlation in 2022", x="Year", y="Infection Rate (%)",  color="Virus", caption="Source: U.S. Centers for Disease Control and Prevention (cdc.gov)") +  theme(aspect.ratio = 2/8) + scale_x_date(date_labels = "%Y", date_breaks = "1 year")

```

